---
title: "Open Policing Project"
author: "Dayton Bossard"
date: "11/14/2019"
output: html_document
---
##Setup 
Install the correct packages and load the libraries in.
If you have not installed the `tidyverse` package, make sure you uncomment the below lines.

```{r setup, message=F, warning=F}
#install.packages("tidyverse") 
#install.packages("lubridate")
library(tidyverse)
library(lubridate)

#load the data
wichita <- read_csv("https://datajournalism.tech/wp-content/uploads/2019/10/wichita.csv")


population <- tibble( subject_race=c("asian/pacific islander", "black","hispanic","other/unknown","white"), num_people =c(19262, 42679, 63659, 13451, 246343))
```

## Data Analysis
Explore the dataset provided by Stanford University. See more on their website
 <https://openpolicing.stanford.edu>.
 
```{r discover-the-dataset}
View(wichita) #to view the data table
str(wichita)  #to see the characteristics of variables
glimpse (wichita) #to see a short summary of values in each column
colnames(wichita) #to view column headers

```
After viewing the dataset, you can analyze it to see the min, max, mean, median and other values for each variable. These are called descriptive statistics.

```{r run-basic-descriptive-statistics}
summary(wichita)
```

There are some verbs that you need to memorize. See more at <https://learn.r-journalism.com/en/wrangling/dplyr/dplyr/>
First, the `select` verb helps you grab column(s) in a dataset

```{r select-one-or-more-columns}
race<- select(wichita,subject_race) #to select the subject_race columns

```

Second, the `group_by` verb helps you categorize your values into fewer groups. The `summarize` verb always goes along with the `group_by` to help count the number of values for each group and compute the percentage of each group over the whole population.

```{r calculate-statistics-for-each-level-of-a-column }
race <- group_by(race, subject_race) %>% summarize(value=n(),  prop=value/nrow(.))

View(race) # to view the `race` table

```


```{r join-two-tables}
stops<-left_join(race, population, by = "subject_race")
```

```{r}
stops <- mutate(stops,stop_rate=value/num_people)
```


##Data Visualization
We will need certain packages to be installed and called before creating our charts.

###Bar Chart
```{r bar-chart}

#make the plot

bar <- ggplot(stops,
       aes(x=reorder(subject_race,stop_rate), y=stop_rate))+
  geom_bar(stat="identity", 
           position="identity", 
           fill="yellow")+
  geom_hline(yintercept = 0) +
  labs(title="Stopped Drivers by Race",
       subtitle = "African American drivers got stopped the most in the city of Wichita Kansas")+
  coord_flip()

options(scipen=10000)

bar
```

###Interactive Map with leaflet
```{r}
#install.packages("httpuv")
#install.packages("leaflet")
library(httpuv)
library(leaflet)
m <- leaflet() %>% 
  addTiles() %>% 
  setView(lng=  -97.317163, lat= 37.685327, zoom=16) %>% 
  addMarkers(lng= -97.317163, lat= 37.685327,  popup="Wichita, KS")

m
```


```{r}
race <- colorFactor(c("coral1", "black", "yellow", "darkolivegreen", "darkgrey"), domain=c("white", "black", "asian/pacific islander", "hispanic", "other/unknown"), ordered=TRUE)
m2<- leaflet(wichita) %>%
  addProviderTiles(providers$OpenStreetMap) %>% 
  setView(lng=-97.31716337, lat= 37.685327, zoom=11) %>% 
  addCircleMarkers(~lng, ~lat, popup=paste("This is a", wichita$subject_race, "and", wichita$subject_sex, "driver."), weight= 1, radius=2, color=~race(subject_race), stroke=F, fillOpacity=1)

m2
```




